---
title: "ANOVA"
output: html_document
---

### Question 1
On the ALL data set, we consider the ANOVA on the gene with the probe “38555_at” expression values on two factors. The first factor is the disease stages: B1, B2, B3 and B4 (we only take patients from those four stages). The second factor is the gender of the patient (stored in the variable ALL$sex).\
(a) Conduct the appropriate ANOVA analysis. Does any of the two factors affects the gene expression values? Are there interaction between the two factors?\
(b)Check the ANOVA model assumption with diagnostic tests? Are any of the assumptions violated?\

```{r warning=FALSE, message=FALSE}
library(ALL)
data(ALL)
library(lmtest)

# (a)
ALLBg <- ALL[,which(ALL$BT %in% c("B1","B2","B3","B4") & ALL$sex %in% c("M","F"))]
y_1 <- exprs(ALLBg)["38555_at",]
Bcell<-ALLBg$BT
gender <- ALLBg$sex
anova(lm(y_1~Bcell*gender))
cat("B stages do affect the expression values, but gender doesn't affect, 
    B stages and genders combined factors doesn't affect expression values either")
anova(lm(y_1~Bcell+gender))
cat("From the test result, we conclude the same as above.")
summary(lm(y_1~Bcell+gender))
cat("Here we can see that all the B stages are have extremely small p-values, 
    but genders p-value is greater than 0.05. Therefore, we can agree the conclusion above.")


# (b)
shapiro.test(residuals(lm(y_1~Bcell*gender)))
bptest(lm(y_1~Bcell*gender))
cat("Shapiro test and bptest show the p-values are identical, 
    so we agree with the assumption from part a")
```

### Question 2
We wish to conduct a permutation test for ANOVA on $(y{_1},...,y{_N})$,with the group\
identifiers stored in the vector ‘group’. We wish to use $\dfrac{1}{(1-g)}\sum_{j=1}^g(\hat \mu{_j}-\hat \mu)^2$ as the test statistic. \
Here $\hat \mu{_j}$ is the j-th group sample mean, and $\hat \mu=\dfrac{1}{g}\sum_{j=1}^g\hat \mu{_j}$ \
(a) Program this permutation test in R.\
(b) Run this permutation test on the Ets2 repressor gene 1242_at on the patients
in stage B1, B2, and B3 from the ALL data set.

```{r warning=FALSE, message=FALSE}
ALLB123 <- ALL[,ALL$BT %in% c("B1","B2","B3")]
y_2 <- exprs(ALLB123)["1242_at",]
group <- ALLB123$BT[,drop=T]
n <- length(y_2)
group.means <- by(y_2, group, mean)
n.group <- length(group.means)
n.perm <- 2000
T.obs <- (1/(n.group))*sum((group.means-mean(group.means))^2)
T.perm <- rep(NA, n.perm)
for(i in 1:n.perm){
  y_2.perm = sample(y_2, n, replace=F) 
  T.perm[i] <- (1/(n.group))*sum((group.means-mean(group.means))^2)
}
mean(T.perm>=T.obs)
cat("p-value > 0.05, so there is no significant difference among 1242_at")

```
