---
title: "Hypothesis Testing"
output: html_document
---

### Question 1
On the Golub et al. (1999) data, consider the “H4/j gene” gene (row 2972) and the “APS Prostate specific antigen” gene (row 2989). Setup the appropriate hypothesis for proving the following claims. Chose and carry out the appropriate tests.\
(a) The mean “H4/j gene” gene expression value in the ALL group is greater than -0.9.\
(b) The mean “H4/j gene” gene expression value in ALL group differs from the mean “H4/j gene” gene expression value in the AML group.\
(c) In the ALL group, the mean expression value for the “H4/j gene” gene is lower than the mean expression value for the “APS Prostate specific antigen” gene.\
(d) Let $p{_{low}}$ denote the proportion of patients for whom the “H4/j gene” expression is lower than the “APS Prostate specific antigen” expression. We wish to show that $p{_{low}}$ in the ALL group is greater than half. Does this test conclusion agree with the conclusion in part (c)?\
(e) Let $p{_{H4j}}$ denotes the proportion of patients for whom the “H4/j gene” expression values is greater than -0.6. We wish to show that $p{_{H4j}}$ in the ALL group is less than 0.5.\
(f) The proportion $p{_{H4j}}$ in the ALL group differs from the proportion $p{_{H4j}}$ in the AML group.\

```{r warning=FALSE, message=FALSE}
library(multtest)
data("golub")
gol.fac <- factor(golub.cl, levels=0:1, labels = c("ALL","AML"))
dataH4j <- golub[2972,]
dataAPS <- golub[2989,]
dataH4j_ALL <- golub[2972, gol.fac=="ALL"]
dataH4j_AML <- golub[2972, gol.fac=="AML"]
dataAPS_ALL <- golub[2989, gol.fac=="ALL"]
dataAPS_AML <- golub[2989, gol.fac=="AML"]
```

### (a) We state $H{_0}$: $\mu=-0.9$ and $H{_A}$: $\mu>-0.9$
```{r warning=FALSE, message=FALSE}
t.test(dataH4j_ALL, alternative = "greater", mu = -0.9)
```
### (b) We state $H{_0}$: $\mu{_{ALL}}=\mu{_{AML}}$ and $H{_A}$: $\mu{_{ALL}}\neq\mu{_{AML}}$
```{r warning=FALSE, message=FALSE}
t.test(dataH4j~gol.fac)
```
### (c) We state $H{_0}$: $\mu{_{H4j}}=\mu{_{APS}}$ and $H{_A}$: $\mu{_{H4j}}<\mu{_{APS}}$
```{r warning=FALSE, message=FALSE}
t.test(dataH4j_ALL, dataAPS_ALL, alternative = "less", mu = 0)
```

### (d) We state $H{_0}$: $p{_{low}}=1/2$ and $H{_A}$: $p{_{low}}>1/2$
```{r warning=FALSE, message=FALSE}
xd <- sum((dataH4j_ALL-dataAPS_ALL)<0)
binom.test(x=xd,n=27,p=0.5,alternative = "greater")
```

### (e) We state $H{_0}$: $p{_{H4j}}=0.5$ and $H{_A}$: $p{_{H4j}}<0.5$
```{r warning=FALSE, message=FALSE}
xe <- sum(dataH4j_ALL>-0.6)
binom.test(x=xe,n=27,p=0.5,alternative = "less")
```

### (f) We state $H{_0}$: $p{_{ALL}}=p{_{AML}}$ and $H{_A}$: $p{_{ALL}}\neq p{_{AML}}$
```{r warning=FALSE, message=FALSE}
x_all <- sum(dataH4j_ALL>-0.6)
x_aml <- sum(dataH4j_AML>-0.6)
prop.test(x=c(x_all,x_aml),n=c(27,11),alternative = "two.sided",correct = F)
```

### Question 2
Suppose that the probability to reject a biological hypothesis by the results of a certain experiment is 0.05. This experiment is repeated 2000 times.\
(a) How many rejections do you expect?\
(b) What is the probability of less than 90 rejections?\

```{r warning=FALSE, message=FALSE}
#(a)
ex_rejection <- 2000 * 0.05

#(b) 
pbinom(89, size = 2000, prob = 0.05)

```

### Question 3
For testing $H{_0}:\mu=3\ versus\  H{_A}:\mu>3$, we considers a new α=0.1 level test which rejects when\
$t{_{obs} = (\bar X - 3)/(s/\sqrt{n})}$ falls between $t{_{0.3,n-1}}$ and $t{_{0.4,n-1}}$.\
(a) Use a Monte Carlo simulation to estimate the Type I error rate of this test\
when n=20. Do 10,000 simulation runs of data sets from the $N(\mu=3, \sigma=4)$ . Please show the R script for the\ simulation, and the R outputs for running the script. Provide your numerical estimate for the Type I error rate. Is\ this test valid (that is, is its Type I error rate same as the nominal α=0.1 level)?\
(b) Should we use this new test in practice? Why or why not?\

```{r warning=FALSE, message=FALSE}
#(a)
nsim3 <- 10000
p.value3 <- rep(NA, nsim3)
p.bon3 <- rep(NA, nsim3)
p.fdr3 <- rep(NA, nsim3)
for(i in 1:nsim3){
  data3 <- rnorm(20,3,4)
  p.value3[i] <- t.test(data3, alternative = "greater", mu = 3, conf.level = 0.9)$p.value
  p.bon3[i] <- p.adjust(p=p.value3[i], method = "bonferroni")
  p.fdr3[i] <- p.adjust(p=p.value3[i], method = "fdr")
}
cat("The error rate is: ", errorRate3 <- sum(p.value3<0.1)/nsim3)

#(b)
cat("No, we shouldn't use this test. Because the error rate is too high.")
```