---
title: "Nonparametric Hypothesis Testing"
output: html_document
---

### Question 1
For the Golub et al. (1999) data set, use appropriate Wilcoxon two-sample tests to find the genes whose mean expression values are higher in the ALL group than in the AML group.\
(a) Use FDR adjustments at the 0.05 level. How many genes are expressed higher in the ALL group?\
(b) Find the gene names for the top three genes with smallest p-values. Are they the same three genes with largest\ difference between the means in the ALL group and the AML group?\

```{r warning=FALSE, message=FALSE}
library(multtest)
data("golub")
# (a)
gol.fac <- factor(golub.cl, levels=0:1, labels = c("ALL","AML"))
nGene <- nrow(golub)
p.values1 <- rep(NA, nGene)
for(i in 1:nGene){
  p.values1[i] <- wilcox.test(golub[i,gol.fac=="ALL"],golub[i,gol.fac=="AML"], paired = F, 
                              alternative = "greater")$p.value
}
p.fdr1 <- p.adjust(p=p.values1, method = "fdr")
cat("There are: ",sum(p.fdr1<0.05),"genes are expressed higher in ALL group")

# (b)
# pIndex1 <- order(p.fdr1, decreasing = FALSE)
# for(i in 1:3){
#   cat(golub.gnames[pIndex1[i],2],"\n")
# }

cat(paste0(golub.gnames[order(p.fdr1, decreasing = FALSE)[1:3],2], sep = "\n"))

```

### Question 2
For the Golub et al. (1999) data set, apply the Shapiro-Wilk test of normality to every gene’s expression values in the AML group. How many genes do not pass the test at 0.05 level with FDR adjustment?

```{r warning=FALSE, message=FALSE}
p.values2 <- rep(NA, nGene)
for(i in 1: nGene){
  p.values2[i] <- shapiro.test(golub[i, gol.fac=="AML"])$p.value
}
p.fdr2 <- p.adjust(p=p.values2, method = "fdr")
cat(sum(p.fdr2<0.05)," genens do not pass the test at 0.05 level with FDR adjustment.")
```
