---
title: "Statistical_Inference"
output: html_document
---

### Question 1
A random sample of size 6 from the exp(λ) 
distribution results in observations: 1.636, 0.374, 0.534, 3.015, 0.932, 0.179. 
Find the MLE on this data set in two ways: by numerical optimization of the likelihood and by the analytic formula.

```{r warning=FALSE}
data1 <- c(1.636,0.374,0.534,3.015,0.932,0.179)
# Analytic
nlik <- function(lam) -prod(lam*exp(-lam*data1))
optim(par = 1, nlik)$par
# Numerical 
nloglikEXP1 <- function(lam) -sum(log2(lam*exp(-lam*data1)))
optim(par = 1, nloglikEXP1)$par
```

### Question 2
A random sample of $X{_1} , ..., X{_{53}} $, from the chi-square distribution with m degree 
of freedom, has sample mean $\bar X = 100.8$ and sample standard deviation s = 12.4 \
(a) Find the point estimator of m using the method of moments. \
(b) Find a one-sided 90% lower confidence interval of m. \
```{r warning=FALSE}
#(a)
m_a <- 100.8
#(b)
m_b <- 100.8+qt(c(0.1), df = 53-1)*12.4/sqrt(53)
cat("One sided 90% lower CI of m is (", m_b, ", inf)")
```
