---
title: "Module 6 Homework"
author: "Yanhe Wen"
date: "2/17/2017"
output:
  pdf_document: default
  html_notebook: default
---

##Question 1
We prepare the data by following R scripts:
```{r results='hide', message=FALSE, warning=FALSE}
library(multtest)
data("golub")
gol.fac <- factor(golub.cl, levels=0:1, labels = c("ALL","AML"))
dataH4j <- golub[2972,]
dataAPS <- golub[2989,]
dataH4j_ALL <- golub[2972, gol.fac=="ALL"]
dataH4j_AML <- golub[2972, gol.fac=="AML"]
dataAPS_ALL <- golub[2989, gol.fac=="ALL"]
dataAPS_AML <- golub[2989, gol.fac=="AML"]
```
###(a)
We state $H{_0}$: $\mu=-0.9$ and $H{_A}$: $\mu>-0.9$
```{r}
t.test(dataH4j_ALL, alternative = "greater", mu = -0.9)
```
We reject null hypothesis.

###(b)
We state $H{_0}$: $\mu{_{ALL}}=\mu{_{AML}}$ and $H{_A}$: $\mu{_{ALL}}\neq\mu{_{AML}}$
```{r}
t.test(dataH4j_ALL, dataAPS_AML, var.equal = FALSE)
```
We reject null hypothesis.

###(c)
We state $H{_0}$: $\mu{_{H4j}}=\mu{_{APS}}$ and $H{_A}$: $\mu{_{H4j}}<\mu{_{APS}}$
```{r}
t.test(dataH4j_ALL-dataAPS_ALL, alternative = "less", mu = 0)
```
We reject null hypothesis.

###(d)
We state $H{_0}$: $p{_{low}}=1/2$ and $H{_A}$: $p{_{low}}>1/2$
```{r}
xd <- sum((dataH4j_ALL-dataAPS_ALL)<0)
binom.test(x=xd,n=27,p=0.5,alternative = "greater")
```
We accept null hypothesis, Plow = 0.6296296, so we agree with the conclusion in part (c)

###(e)
We state $H{_0}$: $p{_{H4j}}=0.5$ and $H{_A}$: $p{_{H4j}}<0.5$
```{r}
xe <- sum(dataH4j_ALL>-0.6)
binom.test(x=xe,n=27,p=0.5,alternative = "less")
```
Accept null hypothesis, equals to 0.5

###(f)
We state $H{_0}$: $p{_{ALL}}=p{_{AML}}$ and $H{_A}$: $p{_{ALL}}\neq p{_{AML}}$
```{r}
x_all <- sum(dataH4j_ALL>-0.6)
x_aml <- sum(dataH4j_AML>-0.6)
prop.test(x=c(x_all,x_aml),n=c(27,11),alternative = "two.sided",correct = F)
```
Reject null hypothesis, they are not equal.

##Question 2
###(a)
```{r}
ex_reject <- 2000*0.05
```
We expect the rejections are: 100

###(b)
```{r}
binom.test(x=ex_reject,n=2000,p=90/2000,alternative = "less")
```
The probability is 5%

##Question 3
###(a)
```{r}
nsim3 <- 10000
p.value3 <- rep(NA, nsim3)
p.bon3 <- rep(NA, nsim3)
p.fdr3 <- rep(NA, nsim3)
for(i in 1:nsim3){
  data3 <- rnorm(20,3,4)
  p.value3[i] <- t.test(data3, alternative = "greater", mu = 3, conf.level = 0.9)$p.value
  p.bon3[i] <- p.adjust(p=p.value3[i], method = "bonferroni")
  p.fdr3[i] <- p.adjust(p=p.value3[i], method = "fdr")
}
cat("The error rate is: ", errorRate3 <- sum(p.value3<0.1)/nsim3)
```
The error rate is closed to a=0.1 level, which is 10000*0.1=1000

###(b)
No, we shouldn't. Because the high rate of error.

##Question 4
###(a)
```{r}
p.values4 <- rep(NA, 3051)
for(i in 1:3051){
  p.values4[i] <- t.test(golub[i,]~gol.fac)$p.value
}
p.bon4 <- p.adjust(p=p.values4, method = "bonferroni")
p.fdr4 <- p.adjust(p=p.values4, method = "fdr")
cat("From Bonferron, there are: ", sum(p.bon4<0.05))
cat("From FDR, there are: ", sum(p.fdr4<0.05))
```
###(b)
```{r}
index4b <- which(p.fdr4<0.05)
sortedData4b <- order(p.values4[index4b], decreasing = FALSE)
cat("The top three strongeet differentially expressed genes are: ")
for(i in 1:3){
  cat(golub.gnames[sortedData4b[i],2],"\n")
}
```
